A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-03-01 sections/Metatheory.tex
Several refactors

\begin{lemma}[Insert]\label{lem:insert}
If $\Gm_L\vdash \tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$
, then $\Gm_L, \al, \Gm_R \sto \Om$.
\end{lemma}
\begin{lemma}[Extract]\label{lem:extract}
If $\Gm_L, \al, \Gm_R \sto \Om$
, then $\exists \tau$ s.t. $\Gm_L\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
\end{lemma}

@@

\paragraph{Declarative transfer.}
Figure~\ref{fig:decl:worklist} defines a relation $\Om \rto \Om'$,
which transfers all judgements in the declarative worklists to the
declarative type system. This relation checks that every judgement entry in the worklist
holds using a corresponding conventional declarative judgement. 
The typing contexts of declarative judgements are recovered using an
auxiliary erasure function $\|\Om\|$
\footnote{In the proof script we do not use the erasure function,
for the declarative system and well-formedness judgments automatically fits
the non-erased declarative worklist just as declarative contexts.}.
The erasure function simply drops all judgment entries from the worklist,
keeping only variable and type variable declarations.

@@

To simplify matters, we introduce an intermediate system that is still declarative
in nature, but has no overlap. This intermediate system differs only in a few 
rules from DK's declarative system; these are the changes:
\begin{enumerate}
\item Restrict the shape of $B$ in the rule $\mathtt{\forall L}$ subtyping rule:
$$
\inferrule*[right=$\mathtt{\forall L'}$]
{B \neq \all[b]B' \\ \Psi \vdash \tau \\ \Psi \vdash [\tau/a]A \le B}
{\Psi \vdash \all A \le B}
$$
\item Drop the redundant rule $\mathtt{Decl1I}$,
which can be easily derived by a combination of
$\mathtt{DeclSub}$, $\mathtt{Decl1I{\To}}$ and $\mathtt{{\le}Unit}$:
$$
\inferrule*[right=$\mathtt{DeclSub}$]
{
    \inferrule*[right=$\mathtt{Decl1I{\To}}$]
    {~}
    {\Psi \vdash () \To 1}
    \\
    \inferrule*[right=$\mathtt{{\le}Unit}$]
    {~}
    {\Psi \vdash 1 \le 1}
}
{\Psi \vdash () \Lto 1}
$$
\item Restrict the shapes of $e$ and $A$ in the subsumption rule $\mathtt{DeclSub}$:
$$
\inferrule*[right=$\mathtt{DeclSub'}$]
{e \neq \lam e' \\ A \neq \all A' \\ \Psi \vdash e \To A \\ \Psi \vdash A \le B}
{\Psi \vdash e \Lto B}
$$
\end{enumerate}
The resulting declarative system has no overlapping rules
and more closely resembles the algorithmic system, which contains
the same shape constraints.

@@

Our algorithm is sound with respect to DK's declarative system.
For any worklist $\Gm$ that reduces successfully,
there is a valid instantiation $\Om$ that transfers all judgements successfully to the declarative system.
\begin{theorem}[Soundness]
If \emph{wf }$\Gm$ and $\Gm \redto \nil$, then $\exists \Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$.
\end{theorem}

@@

We prove completeness by induction on the derivation of $\Om\redto\nil$
and use the non-overlapping declarative system.
Since the declarative worklist is reduced judgment by judgment
(shown in Figure~\ref{fig:decl:worklist}),
the induction always analyses the first judgment until it is satisfied.
As the algorithmic system introduces existential variables,
a declarative rule may correspond to multiple algorithmic rules,
and thus we analyse each of the possible cases.
% \jimmy{Cite DK's proof as the sketch.}
% I later think it might not be needed; the induction is not that hard to think of

Most cases are relatively easy to prove.
The Insert and Extract lemmas are applied when the algorithm uses existential variables,
but transferred to a monotype for the declarative system,
such as Rules 6, 10, 11, 12-17, 21 and 29.
Algorithmic Rules 10 and 11 require special treatment.
When the induction reaches the $\mathtt{{\le}{\to}}$ case,
the first judgment is of shape $A_1 \to A_2 \le B_1 \to B_2$.
One of the corresponding algorithmic judgments is $\al \le A \to B$.
However, the case analysis does not imply that $\al$ is fresh in $A$ and $B$,
therefore Rule~10 cannot be applied and the proof gets stuck.
The following lemma helps us out in those cases:
the success in declarative subtyping indicates the freshness of $\al$ in $A$ and $B$
in its corresponding algorithmic judgment.
In other words, the declarative system does not accept infinite types.
A symmetric lemma holds for $A\to B \le \al$.
%thus, as an example, we cannot find a monotype $\tau$ such that $\tau\le 1\to \tau$,
%which could be transferred by $\al\le 1\to\al$.

@@

Finally, we show that our algorithm terminates. In other words:
\begin{theorem}[Termination]
Given a well-formed input $\Gm$, the algorithm terminates.
\end{theorem}
Our termination proof is based on a lexicographic group of induction measures\\
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
on the worklist $\Gm$. The definitions of these measures can be found in Figure~\ref{fig:measures}. The first two, 
$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$, measure the total size of terms
and the total difficulty of judgments, respectively. In the latter, check judgements
count for 2, inference judgments for 1 and function inference judgments for 3.
The next two measures, $|\cdot|_\forall$ and $|\cdot|_\to$, count the total number of
universal quantifiers and function types, respectively. Finally,
$|\cdot|_{\al}$ counts the number of existential variables in the worklist.

@@

We now show that entirely reducing these smaller problems leaves the worklist
in a state with an overall smaller measure. Our starting point is a 
worklist of the form $\Gm,\Gm_i$ where $\Gm_i$ are instantiation judgments.
\begin{gather*}
\begin{aligned}
\Gm_i &:= \nil \mid \Gm_i, \al\le A \mid \Gm_i, A\le\al \quad
    \text{where } \al\notin FV(A) \cup FV(\Gm_i)
\end{aligned}
\end{gather*}
Fully reducing these instantiation judgments at the top of the worklist
has a twofold impact. Firstly, new entries may be pushed onto the worklist which
are not instantiation judgments. This only happens when $\Gm_i$ contains a universal quantifier
that is reduced by Rule 8 or 9. The new entries then are of the form $\Gm_\le$:
\begin{gather*}
\begin{aligned}
\Gm_\le &:= \nil \mid \Gm_\le, a \mid \Gm_\le, \al \mid \Gm_\le, A\le B
\end{aligned}
\end{gather*}
Secondly, reducing the instantiation judgments may also affect the remainder of the worklist $\Gm$,
by solving existing existentials and introducing new ones. This worklist update is captured in the 
update judgement $\Gm \jExt \Gm'$ defined in Figure~\ref{fig:worklist_ext}.
For instance, existential variable instantiation, 
i.e. $\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
can be derived as a combination of the three rules that define the update relation.

@@

The good news is that worklist update does not affect the three main worklist measures:
\newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}\label{lemma:inst:invariant}
\begin{lemma}[Measure Invariants of Worklist Extension]
If $\Gm\jExt\Gm'$ then $\equivGm{e}$ and $\equivGm\Leftrightarrow$ and $\equivGm\forall$.
\end{lemma}

Moreover, we can characterise the reduction of the instantiation judgments as follows.
\begin{lemma}[Instantiation Decidability]\label{lemma:inst:decidable}
For any well-formed algorithmic worklist $(\Gm, \Gm_i)$,
\begin{enumerate}[1)]
    \item When $|\Gm_i|_\forall = 0$, $(\Gm, \Gm_i) \redto \Gm'$,
        where $|\Gm'|_{\al} = |\Gm|_{\al} - |\Gm_i|$ and $\Gm\jExt\Gm'$.
    \item When $|\Gm_i|_\forall > 0$, $(\Gm, \Gm_i) \redto (\Gm', \Gm_\le)$,
        where $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\jExt\Gm'$.
\end{enumerate}
\end{lemma}
Hence, reducing the instantiation judgment prefix $\Gm_i$ either
decreases the number of universal quantifiers or 
eliminates one existential variable per instantiation judgment.
The proof of this lemma proceeds by induction on the measure $2*|\Gm_i|_\to + |\Gm_i|$
of the instantiation judgment list $\Gm_i$.

Let us go back to the whole algorithm and summarize our findings.
The termination theorem is shown through a lexicographic group of induction measures
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
which is decreased by nearly all rules.
In the exceptional case that the measure does not decrease immediately, 
we encounter an instantiation judgment at the top of the worklist. We can then
make use of Lemma~\ref{lemma:inst:decidable} to show that $|\Gm|_{\al}$ or $|\Gm|_\forall$ decreases
when that instantiation judgment is consumed. Moreover, Lemma~\ref{lemma:inst:invariant} establishes
that no higher-priority measure component increases. Hence, in the exceptional case we have an overall measure decrease too.

@@

\begin{figure}[ht]
\centering\begin{tabular}{|c|c|c|l|}\hline
    File(s) & SLOC & \#Theorems & Description\\\hline
    olist.thm, nat.thm  &  309 & 56  & Basic data structures\\\hline
    typing.thm          &  238 & 5   & Declarative \& algorithmic system, debug examples\\\hline
    decl.thm            &  226 & 33  & Basic declarative properties\\\hline
    order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strenthening\\\hline
    alg.thm             &  655 & 72  & Basic algorithmic properties\\\hline
    trans.thm           &  628 & 53  & \makecell[l]{Transfer and its basic properties \\
                            Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
    declTyping.thm      &  944 & 72  & \makecell[l]{Non-overlapping declarative system \\
                            Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll}, \ref{lem:inv_chkLam}}\\\hline
    soundness.thm       & 1086 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
    depth.thm           &  206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\\hline
    dcl.thm             &  383 & 12  & Non-overlapping declarative worklist \\\hline
    completeness.thm    & 1075 & 61  &
                            Completeness theorem; aux. lemmas and relations\\\hline
    inst\_decidable.thm &  830 & 42  & Other measures; Lemma~\ref{lemma:inst:decidable}\\\hline
    decidability.thm    &  885 & 50  & Decidability theorem\\\hline
    Total               & 7700 & 567 & (51 definitions in total)\\\hline
\end{tabular}
\caption{Statistics for the proof scripts}
\label{fig:proof_statistics}
\end{figure}

>>>>

\begin{lemma}[Insert]\label{lem:insert}
If $\Gm_L, [\tau/\al]\Gm_R \sto \Om$ and $\Gm_L\vdash \tau$
, then $\Gm_L, \al, \Gm_R \sto \Om$.
\end{lemma}
\begin{lemma}[Extract]\label{lem:extract}
If $\Gm_L, \al, \Gm_R \sto \Om$
, then there exists $\tau$ s.t. $\Gm_L\vdash\tau$ and $\Gm_L, [\tau/\al]\Gm_R \sto \Om$.
\end{lemma}

@@

\paragraph{Declarative transfer.}
Figure~\ref{fig:decl:worklist} defines a relation $\Om \rto \Om'$,
which transfers all judgements in the declarative worklists to the
declarative type system. This relation checks that every judgement entry in the worklist
holds using a corresponding conventional declarative judgement. 
The typing contexts of declarative judgements are recovered using an
auxiliary erasure function $\|\Om\|$
\footnote{In the proof script we do not use the erasure function,
for the declarative system and well-formedness judgments automatically fit
the non-erased declarative worklist just as declarative contexts.}.
The erasure function simply drops all judgment entries from the worklist,
keeping only variable and type variable declarations.

@@

To simplify matters, we introduce an intermediate system that is still declarative
in nature, but has no overlap. This intermediate system differs only in a few 
rules from DK's declarative system; these are the changes:
\begin{enumerate}
\item Restrict the shape of $B$ in the rule $\mathtt{\forall L}$ subtyping rule:
$$
\inferrule*[right=$\mathtt{\forall L'}$]
{B \neq \all[b]B' \\ \Psi \vdash \tau \\ \Psi \vdash [\tau/a]A \le B}
{\Psi \vdash \all A \le B}
$$
\item Drop the redundant rule $\mathtt{Decl1I}$,
which can be easily derived by a combination of
$\mathtt{DeclSub}$, $\mathtt{Decl1I{\To}}$ and $\mathtt{{\le}Unit}$:
$$
\inferrule*[right=$\mathtt{DeclSub}$]
{
    \inferrule*[right=$\mathtt{Decl1I{\To}}$]
    {~}
    {\Psi \vdash () \To 1}
    \\
    \inferrule*[right=$\mathtt{{\le}Unit}$]
    {~}
    {\Psi \vdash 1 \le 1}
}
{\Psi \vdash () \Lto 1}
$$
\item Restrict the shapes of $e$ and $A$ in the subsumption rule $\mathtt{DeclSub}$:
$$
\inferrule*[right=$\mathtt{DeclSub'}$]
{e \neq \lam e' \\ A \neq \all A' \\ \Psi \vdash e \To A \\ \Psi \vdash A \le B}
{\Psi \vdash e \Lto B}
$$
\end{enumerate}
The resulting declarative system has no overlapping rules
and more closely resembles the algorithmic system,
which contains constraints of the same shape.

@@

Our algorithm is sound with respect to DK's declarative system.
For any worklist $\Gm$ that reduces successfully,
there is a valid instantiation $\Om$ that transfers all judgements successfully to the declarative system.
\begin{theorem}[Soundness]
If \emph{wf }$\Gm$ and $\Gm \redto \nil$,
then there exists $\Om$ s.t. $\Gm\sto\Om$ and $\Om\redto\nil$.
\end{theorem}

@@

We prove completeness by induction on the derivation of $\Om\redto\nil$
and use the non-overlapping declarative system.
Since the declarative worklist is reduced judgment by judgment
(shown in Figure~\ref{fig:decl:worklist}),
the induction always analyses the first judgment by a small step.
As the algorithmic system introduces existential variables,
a declarative rule may correspond to multiple algorithmic rules,
and thus we analyse each of the possible cases.
% \jimmy{Cite DK's proof as the sketch.}
% I later think it might not be needed; the induction is not that hard to think of

Most cases are relatively easy to prove.
The Insert and Extract lemmas are applied when the algorithm uses existential variables,
but transferred to a monotype for the declarative system,
such as Rules 6, 8, 10, 11, 12-17, 21, 25, 27 and 29.

Algorithmic Rules 10 and 11 require special treatment.
When the induction reaches the $\mathtt{{\le}{\to}}$ case,
the first judgment is of shape $A_1 \to A_2 \le B_1 \to B_2$.
One of the corresponding algorithmic judgments is $\al \le A \to B$.
However, the case analysis does not imply that $\al$ is fresh in $A$ and $B$,
therefore Rule~10 cannot be applied and the proof gets stuck.
The following lemma helps us out in those cases:
the success in declarative subtyping indicates the freshness of $\al$ in $A$ and $B$
in its corresponding algorithmic judgment.
In other words, the declarative system does not accept infinite types.
A symmetric lemma holds for $A\to B \le \al$.
%thus, as an example, we cannot find a monotype $\tau$ such that $\tau\le 1\to \tau$,
%which could be transferred by $\al\le 1\to\al$.

@@

Finally, we show that our algorithm terminates. In other words:
\begin{theorem}[Termination]
Given a well-formed input $\Gm$, the algorithm terminates.
\end{theorem}
Our termination proof is based on a lexicographic group of induction measures\\
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
on the worklist $\Gm$. The definitions of these measures can be found in Figure~\ref{fig:measures}. The first two, 
$|\cdot|_e$ and $|\cdot|_\Leftrightarrow$, measure the total size of terms
and the total difficulty of judgments, respectively. In the latter, check judgements
count for 2, inference judgments for 1 and function inference judgments for 3.
Other two measures, $|\cdot|_\forall$ and $|\cdot|_\to$, count the total number of
universal quantifiers and function types, respectively. Finally,
$|\cdot|_{\al}$ counts the number of existential variables in the worklist,
and $|\cdot|$ is simply the length of the worklist.

@@

We now show that entirely reducing these smaller problems leaves the worklist
in a state with an overall smaller measure. Our starting point is a 
worklist of the form $\Gm,\Gm_i$ where $\Gm_i$ are instantiation judgments.
\begin{gather*}
\begin{aligned}
\Gm_i &:= \nil \mid \Gm_i, \al\le A \mid \Gm_i, A\le\al \quad
    \text{where } \al\notin FV(A) \cup FV(\Gm_i)
\end{aligned}
\end{gather*}
Fully reducing these instantiation judgments at the top of the worklist
has a twofold impact. Firstly, new entries may be pushed onto the worklist which
are not instantiation judgments. This only happens when $\Gm_i$ contains a universal quantifier
that is reduced by Rule 8 or 9. The new entries then are of the form $\Gm_\le$:
\begin{gather*}
\begin{aligned}
\Gm_\le &:= \nil \mid \Gm_\le, a \mid \Gm_\le, \al \mid \Gm_\le, A\le B
\end{aligned}
\end{gather*}
Secondly, reducing the instantiation judgments may also affect the remainder of the worklist $\Gm$,
by solving existing existentials and introducing new ones. This worklist update is captured in the 
update judgement $\Gm \jExt \Gm'$ defined in Figure~\ref{fig:worklist_ext}.
For instance, an existential variable instantiation, 
$\Gm_L,\al,\Gm_R \jExt \Gm_L, \al[1], \al[2], [\al[1] \to \al[2] / \al]\Gm_R$,
can be derived as a combination of the three rules that define the update relation.

@@

The good news is that worklist update does not affect the three main worklist measures:
\newcommand{\equivGm}[1]{|\Gm|_{#1} = |\Gm'|_{#1}}
\begin{lemma}[Measure Invariants of Worklist Extension]\label{lemma:inst:invariant}
If $\Gm\jExt\Gm'$ then $\equivGm{e}$ and $\equivGm\Leftrightarrow$ and $\equivGm\forall$.
\end{lemma}

Moreover, we can characterise the reduction of the instantiation judgments as follows.
\begin{lemma}[Instantiation Decidability]\label{lemma:inst:decidable}
For any well-formed algorithmic worklist $(\Gm, \Gm_i)$:
\begin{enumerate}[1)]
    \item If $|\Gm_i|_\forall = 0$,
        then there exists $\Gm'$\\
        s.t. $(\Gm, \Gm_i) \redto \Gm'$ and $|\Gm'|_{\al} = |\Gm|_{\al} - |\Gm_i|$ and $\Gm\jExt\Gm'$.
    \item If $|\Gm_i|_\forall > 0$,
        then there exist $\Gm', \Gm_\le$\\
        s.t. $(\Gm, \Gm_i) \redto (\Gm', \Gm_\le)$ and $|\Gm_\le|_\forall = |\Gm_i|_\forall - 1$ and $\Gm\jExt\Gm'$.
\end{enumerate}
\end{lemma}
Hence, reducing the instantiation judgment prefix $\Gm_i$ either
decreases the number of universal quantifiers or 
eliminates one existential variable per instantiation judgment.
The proof of this lemma proceeds by induction on the measure $2*|\Gm_i|_\to + |\Gm_i|$
of the instantiation judgment list $\Gm_i$.

Let us go back to the whole algorithm and summarize our findings.
The termination theorem is shown through a lexicographic group of induction measures
$\langle |\Gm|_e, |\Gm|_\Leftrightarrow, |\Gm|_\forall, |\Gm|_{\al}, |\Gm|_\to + |\Gm| \rangle$
which is decreased by nearly all rules.
In the exceptional case that the measure does not decrease immediately, 
we encounter an instantiation judgment at the top of the worklist. We can then
make use of Lemma~\ref{lemma:inst:decidable} to show that $|\Gm|_{\al}$ or $|\Gm|_\forall$ decreases
when that instantiation judgment is consumed or partially reduced.
Moreover, Lemma~\ref{lemma:inst:invariant} establishes
that no higher-priority measure component increases.
Hence, in the exceptional case we have an overall measure decrease too.

@@

\begin{figure}[ht]
\centering\begin{tabular}{|c|c|c|l|}\hline
    File(s) & SLOC & \#Theorems & Description\\\hline
    olist.thm, nat.thm  &  309 & 56  & Basic data structures\\\hline
    typing.thm          &  238 & 5   & Declarative \& algorithmic system, debug examples\\\hline
    decl.thm            &  226 & 33  & Basic declarative properties\\\hline
    order.thm           &  235 & 27  & The $|\cdot|_\forall$ measure; decl. subtyping strenthening\\\hline
    alg.thm             &  655 & 72  & Basic algorithmic properties\\\hline
    trans.thm           &  628 & 53  & \makecell[l]{Worklist instantiation with basic properties \\
                            Lemmas~\ref{lem:insert}, \ref{lem:extract}}\\\hline
    declTyping.thm      &  944 & 72  & \makecell[l]{Non-overlapping declarative system \\
                            Lemmas~\ref{lem:inv_allR}, \ref{lem:inv_chkAll}, \ref{lem:inv_chkLam}}\\\hline
    soundness.thm       & 1086 & 70  & Soundness theorem; aux. lemmas on transfer\\\hline
    depth.thm           &  206 & 14  & The $|\cdot|_\to$ measure; Lemma~\ref{lem:prune_inst}\\\hline
    dcl.thm             &  383 & 12  & Non-overlapping declarative worklist \\\hline
    completeness.thm    & 1075 & 61  &
                            Completeness theorem; aux. lemmas and relations\\\hline
    inst\_decidable.thm &  830 & 42  & Other measures; Lemma~\ref{lemma:inst:decidable}\\\hline
    decidability.thm    &  885 & 50  & Decidability theorem\\\hline
    \hline\emph{Total}               & 7700 & 567 & (51 definitions in total)\\\hline
\end{tabular}
\caption{Statistics for the proof scripts}
\label{fig:proof_statistics}
\end{figure}

### 2019-03-01 sections/algorithm.tex
Several refactors

\begin{figure}
\begin{gather*}
\begin{aligned}
% \text{Type variables}\qquad&a, b\\
\text{Existential variables}\qquad&\al, \bt
\\
%\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
% \text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
%\text{Context}\qquad&\Psi &::=&\quad \nil \mid \Psi, a \mid \Psi, x:A' \color{red} \\[2mm]
% \text{Declarative Judgements}\qquad&\jo &::=&\quad \nil \mid \P\vdash A' \le B' \jc \jo\\
% &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
% \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[2mm]
\text{Algorithmic types}\qquad&A, B, C &::=& \quad 1 \mid a \mid \forall a. A \mid A\to B \mid \al % \quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
\\
% \text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
% \\[2mm]
\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
\\
\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg%\\
% \text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
\end{aligned}
\end{gather*}

@@

Rules 22-26 deal with type inference judgments, written in continuation-passing-style.
When an inference judgment succeeds with type $A$,
the algorithm continues to work on the inner-chain $\jg$ by
assigning $A$ to its placeholder variable $a$.
Rule 23 infers an annotated expression by changing into checking mode,
therefore another judgment chain is created.
Rule 24 is a base case,
where the unit type $1$ is inferred and thus passed to its child judgment chain.
Rule 26 infers the type of an application by
firstly inferring the type of the function $e_1$,
and then leaving the rest work to an application inference judgment,
which passes $a$, representing the return type of the application,
to the reminder of the judgment chain $\jg$.

>>>>

\begin{figure}
\begin{gather*}
\begin{aligned}
% \text{Type variables}\qquad&a, b\\
\text{Existential variables}\qquad&\al, \bt
\\
%\text{Types}\qquad&A', B', C' &::=&\quad 1 \mid a \mid \forall x. A' \mid A'\to B'\\
% \text{Mono-types}\qquad&\tau &::=&\quad 1 \mid a \mid \tau_1\to \tau_2\\
%\text{Context}\qquad&\Psi &::=&\quad \nil \mid \Psi, a \mid \Psi, x:A' \color{red} \\[2mm]
% \text{Declarative Judgements}\qquad&\jo &::=&\quad \nil \mid \P\vdash A' \le B' \jc \jo\\
% &&&\quad \mid \P\vdash e\Leftarrow A' \jc \jo \mid \P\vdash e\To A' \jc \jo \mid \P\vdash A' \bullet e\TTo C' \jc \jo\\
% \noalign{\jimmy{\text{What do the markers do: $\G \vdash,a \vdash,b\vdash \j_1;\j_2;\j_3$ means $\G,a,b\vdash \j_1 \land \G,a\vdash \j_1,\j_2 \land \G\vdash \j_3$}}} \\[2mm]
\text{Algorithmic types}\qquad&A, B, C &::=& \quad 1 \mid a \mid \all A \mid A\to B \mid \al % \quad 1 \mid a \mid \al \mid \forall a. A \mid A\to B
\\
% \text{Terms}\qquad&e&::=&\quad x \mid () \mid \lam{e} \mid e_1~e_2 \mid (e:A)
% \\[2mm]
\text{Algorithmic judgement chain}\qquad&\jg &::=&\quad A \le B \mid e\Lto A \mid e\To_{a} \jg \mid \appInfAlg{A}{e}
\\
\text{Algorithmic worklist}\qquad&\Gm &::=&\quad \nil \mid \Gm, a \mid \Gm, \al \mid \Gm, x: A \mid \Gm \Vdash \jg%\\
% \text{Declarative worklist}\qquad&\Om &::=&\quad \nil \mid \Om, a \mid \Om, x: A \mid \Om \Vdash \jg
\end{aligned}
\end{gather*}

@@

Rules 22-26 deal with type inference judgments, written in continuation-passing-style.
When an inference judgment succeeds with type $A$,
the algorithm continues to work on the inner-chain $\jg$ by
assigning $A$ to its placeholder variable $a$.
Rule 23 infers an annotated expression by changing into checking mode,
therefore another judgment chain is created.
Rule 24 is a base case,
where the unit type $1$ is inferred and thus passed to its child judgment chain.
Rule 26 infers the type of an application by
firstly inferring the type of the function $e_1$,
and then leaving the rest work to an application inference judgment,
which passes $a$, representing the return type of the application,
to the remainder of the judgment chain $\jg$.
